### 2015-08-01 lucas.tex
many changes up to apge4

Our survey collected data on software engineers' views of commonly held software engineering ``laws''.
One of the laws questioned relates directly to the problems associated with
the delayed issue effect: ``Requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' (from Glass~\cite{glass02} p.71 who references Boehm \& Basili~\cite{boehm01}). We abbreviate this law as RqtsErr
(and is
illustrated in \fig{b81}).

The survey was conducted in two phases using Amazon's Mechanical Turk. The first phase was conducted only with professional software engineers solicited through the Mechanical Turk,
Professional software engineers were required to complete a pretest to verify their status as a professional or open source software developer and to confirm their knowledge of basic software engineering terminology and technology. The  second survey was conducted Program Committee members of the ESEC/FSE 2015 and ICSE 2014  conferences solicited via email.

The respondents answered questions on two scales: \newline
\textbf{Agreement:} ``Based on your experience, do you agree that the statement above is correct?'' A Likert scale captured the agreement score from Strongly Disagree to Strongly Agree. A text box was provided to explain the answer. \newline
\textbf{Applicability:} ``To the extent that you believe it, how widely do you think it applies among software development contexts?'' The possible answers were: 
-1 meaning ``I don't know''; and 0 meaning ``this law does not apply at all'';
 1 meaning ``applies in a very narrow range of projects'';  2 meaning
``rarely applies'';   3 meaning
``occasionally applies'';  
4 meaning ``very frequently applies''; and
 5 meaning ``always applies''.
Respondents were required to explain the applicability score in a text box.

In order to baseline the participant's answers, participants were presented with the RqtsErr law and others. For the purposes
of this paper, the nature of the other laws other than RqtsErr are not relevant-- we
only added them in as a way to calibrate responses to the RqtsErr question. All laws were drawn from \cite{glass02} and \cite{endres03}. 
The PC member survey contained an additional question on ``In general, the longer errors are in the system (requirements errors, design errors, coding errors, etc.), the more expensive they are to fix'' (the Delayed Issue Effect, for short). Responses were recorded using the agreement question Likert scale.

Summary statistics for the agreement and applicability scores for the RqtsErr and PhaseDelay laws are presented in Figure~\ref{fig:survey_results}. Responses whose Applicability response was ''I don't know'' are omitted from analysis.

\begin{figure}[!ht] 
\scriptsize

\begin{tabular}{l|c|c|c|c|c}
 &  & \multicolumn{2}{c}{agreement} & \multicolumn{2}{c}{applicability} \\\cline{1-1} 
Practitioner survey  & N & $\bar{x}$ & $\tilde{x}$ & $\bar{x}$ & $\tilde{x}$ \\
\hline 
\textbf{Rqts errors are most expensive...} & 16 & 4.6 & 5 & 4.2 & 4 \\ 
Process maturity improves output & 17 & 4.4 & 4 & 3.9 & 4 \\ 
Most time is spent removing errors & 16 & 4.1 & 4 & 3.9 & 4 \\ 
Missing reqts are hardest to fix & 17 & 4.1 & 4 & 4.0 & 4 \\
Reuse increases prod. and qual. & 16 & 3.9 & 4 & 3.8 & 4 \\
OO-programming reduces errors & 13 & 3.9 & 4 & 3.5 & 4 \\
80-20 rule (defects to modules) & 12 & 3.8 & 4 & 4.0 & 4 \\
Adding manpower to a late project & 15 & 3.7 & 4 & 3.7 & 4 \\
Inspections can remove 90\% of defects & 18 & 3.3 & 4 & 3.5 & 4 \\
Smaller changes have higher error density & 14 & 2.8 & 3 & 3.4 & 3.5 \\
A developer is unsuited to test own code & 17 & 2.6 & 3 & 3.5 & 4\\\hline

Researcher survey \\\hline 
Process maturity improves output & 4 & 4.0 & 4 & 4.0 & 4 \\
\textbf{Rqts errors are most expensive...} & 30 & 3.8 & 4 & 3.8 & 4   \\ 
80-20 rule (defects to modules) & 6 & 3.7 & 4 & 3.5 & 4 \\
\textbf{PhaseDelay} & 30 & 3.6 & 4 & -- & --  \\ 
Missing reqts are hardest to fix & 7 & 3.6 & 4 & 3.7 & 4 \\
Adding manpower to a late project & 4 & 3.5 & 4 & 3.5 & 4 \\
Inspections can remove 90\% of defects & 7 & 3.4 & 4 & 3.6 & 3 \\
Reuse increases prod. and qual. & 6 & 3.3 & 4 & 3.0 & 4 \\
OO-programming reduces errors & 6 & 3.2 & 4 & 2.7 & 3 \\
Most time is spent removing errors & 6 & 2.8 & 3 & 3.2 & 4 \\ 
Smaller changes have higher error density & 4 & 2.8 & 3 & 4.0 & 4 \\
A developer is unsuited to test own code & 7 & 2.1 & 2 & 2.4 & 3
\end{tabular}

\caption{Agreement and applicability of SE axioms.}
\label{fig:survey_results}
\end{figure}

Both  practitioners and researchers  strongly believed in RqtsErr (in both sets of responses, RqtsErr, received  scores higher than most
other laws. Further, in the case of practitioners, this law was rated
as the single most believed effect. 
%The free response texts, most participants agreed that addressing errors %%late may mean system redesign unless the system and the affected %equirement are simple/trivial. Further, any changes to the production %ystem have additional complexity (compared to changes while system in %evelopment) due to users, data existence, and architecture dependencies. 
Caveat: in the free repsonse texts, we note that the researchers who disagreed with the law generally asserted that requirements change can be expensive, but that depends on the process used (e.g., agile vs. waterfall) and the adaptability of the system architecture.

Overall, the RqtsErr law was the most agreed upon and most applicable law of 11 surveyed amongst practitioners, and the second most agreed upon law amongst researchers. The survey results strongly support that the notion that the  delayed issue effect   is a widely-held belief in software engineering.