### 2016-06-15 tex/fig13.tex
Updates from ShareLaTeX

\resizebox{\columnwidth}{!}{\begin{tabular}{c@{~}|l@{~}r@{~}|rrr|r@{~}l|r@{~}l}
  & \multicolumn{2}{c|}{ } & \multicolumn{3}{c|}{Percentiles} & \multicolumn{4}{c}{Growth with respect to earliest phase  } \\ 
   & \multicolumn{2}{c|}{Phase} & \multicolumn{3}{c|}{(units= time in miniutes)} & \multicolumn{4}{c}{(unitless, since these are ratios of two time values) } \\\hline

  rank & injected & removed & 50th & 90th &  IQR & \multicolumn{2}{c|}{50th percentile growth} & \multicolumn{2}{c}{90th percentile growth} \\ 
\hline\multicolumn{10}{c}{~}  \\
1 &  Before   &   DesignInspect   & 10 & 32 & 14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.00 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &   &   CodeReview      & 8 & 31 & 14 & 0.80 & \textcolor{black}{\rule{8mm}{2mm}} & 0.97 & \textcolor{Gray}{\rule{9mm}{2mm}}\\
1 &   &   CodeInspect     & 10 & 32 & 16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.00 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &   &   UnitTest        & 12 & 45 & 21 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}} & 1.41 & \textcolor{Gray}{\rule{14mm}{2mm}}\\
1 &   &   IntTest         & 15 & 63 & 31 & 1.50 & \textcolor{black}{\rule{15mm}{2mm}} & 1.97 & \textcolor{Gray}{\rule{19mm}{2mm}}\\
1 &   &   SysTest         & 11 & 46 & 22 & 1.10 & \textcolor{black}{\rule{12mm}{2mm}} & 1.44 & \textcolor{Gray}{\rule{14mm}{2mm}}\\  
\hline\multicolumn{10}{c}{~}  \\
1 &  Planning     &   ReqtsReview     & 8 & 35 & 14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.00 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &               &   DesignInspect   & 11 & 31 & 13 & 1.38 & \textcolor{black}{\rule{13mm}{2mm}} & 0.89 & \textcolor{Gray}{\rule{8mm}{2mm}}\\
2 &               &   UnitTest        & 24 & 53 & 25 & 3.00 & \textcolor{black}{\rule{30mm}{2mm}} & 1.51 & \textcolor{Gray}{\rule{15mm}{2mm}}\\
\hline\multicolumn{10}{c}{~}  \\
1 &  Reqts   &   ReqtsReview     & 13 & 42 & 20 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.00 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &          &   ReqtsInspect    & 12 & 40 & 18 & 0.92 & \textcolor{black}{\rule{9mm}{2mm}} & 0.95 & \textcolor{Gray}{\rule{9mm}{2mm}}\\
1 &          &   DesignReview    & 10 & 34 & 14 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}} & 0.81 & \textcolor{Gray}{\rule{8mm}{2mm}}\\
1 &          &   DesignInspect   & 9 & 38 & 15 & 0.69 & \textcolor{black}{\rule{6mm}{2mm}} & 0.90 & \textcolor{Gray}{\rule{9mm}{2mm}}\\
1 &          &   CodeInspect     & 13 & 45 & 24 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.07 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &          &   UnitTest        & 10 & 40 & 17 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}} & 0.95 & \textcolor{Gray}{\rule{9mm}{2mm}}\\
1 &          &   IntTest         & 33 & 95 & 42 & 2.54 & \textcolor{black}{\rule{25mm}{2mm}} & 2.26 & \textcolor{Gray}{\rule{22mm}{2mm}}\\
1 &          &   SysTest         & 24 & 126 & 108 & 1.85 & \textcolor{black}{\rule{18mm}{2mm}} & 3.00 & \textcolor{Gray}{\rule{30mm}{2mm}}\\
\hline\multicolumn{10}{c}{~}  \\
1 & Design   &   DesignReview    & 11 & 37 & 16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.00 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &   &   DesignInspect   & 8 & 28 & 12 & 0.73 & \textcolor{black}{\rule{7mm}{2mm}} & 0.76 & \textcolor{Gray}{\rule{7mm}{2mm}}\\
1 &   &   CodeReview      & 10 & 40 & 18 & 0.91 & \textcolor{black}{\rule{9mm}{2mm}} & 1.08 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &   &   CodeInspect     & 9 & 33 & 14 & 0.82 & \textcolor{black}{\rule{8mm}{2mm}} & 0.89 & \textcolor{Gray}{\rule{8mm}{2mm}}\\
1 &   &   UnitTest        & 11 & 41 & 18 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.11 & \textcolor{Gray}{\rule{12mm}{2mm}}\\
1 &   &   IntTest         & 17 & 75 & 31 & 1.55 & \textcolor{black}{\rule{15mm}{2mm}} & 2.03 & \textcolor{Gray}{\rule{20mm}{2mm}}\\
1 &   &   AcceptTest      & 14 & 44 & 19 & 1.27 & \textcolor{black}{\rule{12mm}{2mm}} & 1.19 & \textcolor{Gray}{\rule{12mm}{2mm}}\\
1 &   &   SysTest         & 13 & 40 & 18 & 1.18 & \textcolor{black}{\rule{12mm}{2mm}} & 1.08 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
\hline\multicolumn{10}{c}{~}  \\
1 &  Code   &   CodeReview    & 10 & 35 & 16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 1.00 & \textcolor{Gray}{\rule{10mm}{2mm}}\\
1 &    &   CodeInspect   & 10 & 32 & 15 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} & 0.91 & \textcolor{Gray}{\rule{9mm}{2mm}}\\
1 &   &   UnitTest      & 12 & 45 & 20 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}} & 1.29 & \textcolor{Gray}{\rule{12mm}{2mm}}\\
1 &    &   IntTest       & 14 & 58 & 25 & 1.40 & \textcolor{black}{\rule{14mm}{2mm}} & 1.66 & \textcolor{Gray}{\rule{16mm}{2mm}}\\
1 &    &   AcceptTest    & 16 & 60 & 25 & 1.60 & \textcolor{black}{\rule{16mm}{2mm}} & 1.71 & \textcolor{Gray}{\rule{17mm}{2mm}}\\
1 &   &   SysTest       & 13 & 47 & 20 & 1.30 & \textcolor{black}{\rule{13mm}{2mm}} & 1.34 & \textcolor{Gray}{\rule{13mm}{2mm}}\\

\end{tabular}}